manuscript_figures

### 2024-06-03 scripts/manuscript_figures.Rmd
made figure 4

@@

# Make figure 3 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
merged.dat.morph <- left_join(morph.filt, fish.filt)

# Aggregate the data by year, site, and transect and compute mean values
aggregated.data.morph <- merged.dat.morph %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE)) %>%
  ungroup()

@@

# Extract p-value for mean_kelp_cover
p_value_kelp_cover <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = mean_kelp_biomass, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Kelp Biomass (g/m2)", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

@@

>>>>

# Merge fish and morph data
merged.dat.morph <- left_join(morph.filt, fish.filt)

@@

# Make figure 3 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated.data.morph <- merged.dat.morph %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_kelp_biomass = mean(average_kelp_biomass, na.rm = TRUE)) %>%
  ungroup()

@@

# Extract p-value for mean_kelp_cover
p_value_kelp_biomass <- summary(lmm_model)$coefficients["mean_kelp_biomass", "Pr(>|t|)"]

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = mean_kelp_biomass, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Kelp Biomass (g/m2)", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

@@

# Figure 4: organize rugosity data and make linear regression figure for fish biomass and rugosity
```{r}
# Calculate mean rugosity complexity scores across depths for each transect, grouped by site and year
mean_rugosity <- rugosity %>%
  group_by(SITE, YEAR, TRANSECT) %>%
  summarize(mean_rugosity = mean(c(X0M, X4M, X8M, X12M, X16M, X20M, X24M, X28M, X32M, X36M, X40M), na.rm = TRUE))

# change transect from integer to chr
mean_rugosity$TRANSECT <- as.character(mean_rugosity$TRANSECT)

# Left jon fish and rugosity data
merged.dat.rug <- left_join(mean_rugosity, fish.filt)

# Make figure 4 and run statistical analysis
```{r}
# Aggregate the data by year, site, and transect and compute mean values
aggregated.data.rug <- merged.dat.rug %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarize(mean_fish_biomass = mean(Fish.Biomass..g.m2., na.rm = TRUE),
            mean_rugosity = mean(mean_rugosity, na.rm = TRUE)) %>%
  ungroup()

# Fit the linear mixed effects model
lmm_model <- lmer(mean_fish_biomass ~ mean_rugosity + SITE + (1 | SITE/YEAR), data = aggregated.data.rug)
#check normality
qqPlot(residuals(lmm_model)) #data is normal no need to transform

# Extract p-value for mean_kelp_cover
p_value_rugosity <- summary(lmm_model)$coefficients["mean_rugosity", "Pr(>|t|)"]

# Create scatter plot
figure.4 <- ggplot(aggregated.data.rug, aes(x = mean_rugosity, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_rugosity, 2)), 
           hjust = 2.1, vjust = 1.1, size = 4, parse = FALSE) +
  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Rugosity Complexity") +
  theme_classic();figure.4

# Save the plot to a file
ggsave("output/figures/figure_4.png", plot = figure.4, width = 8, height = 6)
```
